API design for machine learning software: experiences from the scikit-learn project

### 2013-06-26 paper.tex
Cosmit.
Use project URL for sofia-ml and VW.

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML \citep{sculley2010web}
and Vowpal~Wabbit \citep{langford2009sparse}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process inputs/outputs,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.

@@

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power, and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of
the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as
part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
To maximize ease of use, we encourage researchers
to follow these conventions when releasing machine learning software.

>>>>

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/~vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process inputs/outputs,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.

@@

While part of the \sklearn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
The power, and extensibility of the \sklearn API is evidenced
by the large and growing user-base, its use to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the \sklearn conventions. Examples of such packages include
\textit{astroML}\footnote{\url{http://astroml.org/}}
\citep{vanderplas2012astroML}, a package providing
machine learning tools for astronomers and
\textit{wiseRF}\footnote{\url{http://wise.io}}, a commercial random forest
implementation. The source code of
the recently-proposed sparse multiclass
algorithm of \citet{mblondel-mlj2013}, released as
part of the
\textit{lightning}\footnote{\url{https://github.com/mblondel/lightning}}
package, also follows the \sklearn conventions.
To maximize ease of use, we encourage more researchers
to follow these conventions when releasing their software.
